---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Gokila Rajaiah"
date: "March 3, 2018"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


## The Tidyverse Packages

#### 1. 

Below list gives packages for differnt tasks related to data.

  * Plotting - ggplot2
  * Data munging/Wrangling - dplyr
  * Reshaping (gathering/spreading) data - tidyr
  * Importing/Exporting data - readr
  
#### 2.

Functions from all the above task related packages.

  * Plotting - ggplot(), geom_point(), facet_wrap()
  * Data munging/Wrangling - select(), filter(), arrange()
  * Reshaping data - mutate(), gather(), separate()
  * Importing/Exporting data - read_csv(), write_csv(), parse_*() functions.
  
## R Basics

#### 1.

Remove '!' from the variable name. 

```{r}

My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )

```

#### 2.

Fix: 'C' in uppercase changed to lowercase. Added close ' after "it" item. 

```{r}

my_string <- c('has', 'an', 'error', 'in', 'it')

```

#### 3.

All numeric values of the vector have been implicitly converted to string type as 3 & 4 were enclosed in single quotes.
So to perform numeric operations in the vector they have to be converted using cast function.


```{r, error=FALSE, warning=FALSE}

my_vector <- c(1, 2, '3', '4', 5)
my_vector
## my_vector[1] + 3 -- This gives error "non-numeric argument to binary operator"


my_vector1 <- c(1, 2, 3, 4, 5)
my_vector1
my_vector1[1] + 3

```

## Data Import/Export 

#### 1. Import .txt file:

```{r}

file_path <- './Input/rail_trail.txt'

rail_text <- read_delim(file_path, '|')

glimpse(rail_text)


```

#### 2.Export file into R specific Format & Import:

```{r}

saveRDS(rail_text, file = './Output/rail_trail.rds')

rail_textRds <- readRDS('./Output/rail_trail.rds')

glimpse(rail_textRds)


```


## Visualization:

1. Though at a quick first glance the chart seems to be showing that most % of  different group of people are favoring towards Women President in their lifetime, i.e. all left bubbles (Yes) are big compared to the other, this chart is totally wrong. 

  * Title says that below data shows Percentage of respondents. But each of the category breakdown don't add up to 100%.
  * Even if the % values are corrected, this chart doesn't show what was the total Respondents & % of each category of respondents (i.e Age group <45, 45-60, 60+ and Gender type Male/Female).so the chart is not so useful. 
  
2. 

```{r fig.align='center', fig.height=7, fig.width=10}

ggplot(data = diamonds, mapping = aes(x=cut, y=carat, fill=color)) +
  geom_boxplot(position = "identity") +
  coord_flip() +
  ylab("CARAT OF DIAMOND") +
  xlab("CUT OF DIAMOND")

```

3. 

In the above plot we can't really tell the color vs carat under each type of cut. If we change the position to "Dodge", you can tell the range of diamond's weight (Carat) for each color & cut type.   

```{r fig.align='center', fig.cap='CUT Vs Carat by Color', fig.height=7, fig.width=10}

ggplot(data = diamonds, mapping = aes(x=cut, y=carat, fill=color)) +
  geom_boxplot(position = "dodge") +
  coord_flip() +
  ylab("CARAT OF DIAMOND") +
  xlab("CUT OF DIAMOND")


```

## Data munging and wrangling

#### 1. 

In table2 one observation is spread across two rows where one row has the total population and another row shows total cases. In order to tidy this data we need to use spread(). 



```{r}

table2 %>%
  spread(key=type, value=count)

```

#### 2.

```{r}
diamonds %>%
  mutate(price_per_carat = (price/carat))

```

  

#### 3.

```{r}

diamonds

diamonds %>% 
  group_by(cut) %>%
  summarise(num_diamonds = sum(price > 10000.0 & carat < 1.5),
            prop_diamonds = mean(price > 10000 & carat < 1.5)
  )

```



## EDA

#### 1. 

"txhousing" data is from the years between 2000 - 2009 (inclusive).


```{r}

txhousing %>% group_by(year) %>% count()

```

#### 2.

46 citites are represented in the txhousing dataset.

```{r}

city_txhousing <- txhousing %>%
  group_by(city) %>%
  summarise(n())

glimpse(city_txhousing)

```

#### 3.

```{r}

txhousing

```
  




