---
title: "COMPSCIX 415.2 Homework 8"
author: "Gokila Rajaiah"
date: "March 26, 2018"
output: html_document
---

```{r setup, include=FALSE}

library(tidyverse)



```

#### 1. 

There are 891 observations of 12 variables in the titanic training data. 

```{r}

file_path <- './Input/train_hw8.csv'

titanic_data <- read.csv(file_path)

glimpse(titanic_data)


#convert target variable (Survived) to be a factor so that it is not loaded as integer. 

titanic_data <- titanic_data %>%
  mutate(Survived_fct = factor(Survived, ordered = FALSE),
         Pclass_fct = factor(Pclass, ordered = FALSE))

#titanic_data <- titanic_data %>%
#  mutate(survived_fct = factor(case_when(Survived == 0 ~ 'No',
#                                         Survived == 1 ~ 'Yes')))

```

#### 2.

```{r, error=FALSE, message=FALSE}

#prepare training & test set from the data

set.seed(29283)

train_set <- titanic_data %>% sample_frac(0.7)


test_set <- titanic_data %>% filter(!(PassengerId %in% train_set$PassengerId))


```

```{r}

ggplot(data = titanic_data, mapping = aes(x=Age, y=Survived_fct)) +
  geom_point(aes(color=Sex))


ggplot(titanic_data, aes(Pclass, fill=Survived_fct)) +
  geom_histogram(position='dodge')

ggplot(titanic_data, aes(Age, fill=Survived_fct)) +
  geom_histogram(position='dodge')


```


#### 3.

Looking the features Age, Sex & PClass we can see that 
  * % of first class people survived it higher than % of third class people who did not survive is higher. 
  * % of children/infants survived is higher.
  * % of Females survived is higher than Males.
  
Looking at the logistic regression model coeff, 
  * A decrease in Age is associated with increased probability of survival. 
  * log Odds of survival for Male is -2.66.i.e. the odds of survival for Male is exp(-2.66)
  
```{r}

library(broom)

# Fit a model with intercept only
mod_1 <- glm(Survived_fct ~ Age+Pclass+Sex, data = train_set, family = 'binomial')

# take a look at the features and coefficients
tidy(mod_1)

```

#### 4. 

Note: I am getting the following error: Error: all(fitted[["(fitted)"]] %in% nt) is not TRUE when trying to plot the tree using partykit with 3 features in the model. 

```{r, error=FALSE, message=FALSE}

library("rpart")
library("partykit")


tree_mod <- rpart(Survived_fct ~ Sex+Pclass_fct+Age, data = train_set)

plot(tree_mod)
text(tree_mod, use.n = TRUE)

#summary(tree_mod)

#as.party(tree_mod)

#plot(as.party(tree_mod))

```


#### 5.

```{r, error=FALSE, message=FALSE}
test_logit <- predict(mod_1, newdata = test_set, type = 'response')
test_tree <- predict(tree_mod, newdata = test_set)[,2]

#test_logit
#test_tree
```

```{r}

library(ROCR)

# create the prediction objects for both models
pred_logit <- prediction(predictions = test_logit, labels = test_set$Survived)
pred_tree <- prediction(predictions = test_tree, labels = test_set$Survived)


# get the FPR and TPR for the logistic model
# recall that the ROC curve plots the FPR on the x-axis
perf_logit <- performance(pred_logit, measure = 'tpr', x.measure = 'fpr')
perf_logit_tbl <- tibble(perf_logit@x.values[[1]], perf_logit@y.values[[1]])

# Change the names of the columns of the tibble
names(perf_logit_tbl) <- c('fpr', 'tpr')

# get the FPR and TPR for the tree model
perf_tree <- performance(pred_tree, measure = 'tpr', x.measure = 'fpr')
perf_tree_tbl <- tibble(perf_tree@x.values[[1]], perf_tree@y.values[[1]])

# Change the names of the columns of the tibble
names(perf_tree_tbl) <- c('fpr', 'tpr')

# Plotting function for plotting a nice ROC curve using ggplot
plot_roc <- function(perf_tbl) {
  p <- ggplot(data = perf_tbl, aes(x = fpr, y = tpr)) +
  geom_line(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, lty = 3) +
  labs(x = 'False positive rate', y = 'True positive rate') +
  theme_bw()
  return(p)
}


# Create the ROC curves using the function we created above
plot_roc(perf_logit_tbl)
plot_roc(perf_tree_tbl)

```


Looking at the AUC values it looks like Decision tree model is performing a little better than logistic regression model. 

```{r}

auc_logit <- performance(pred_logit, measure = 'auc')
auc_tree <- performance(pred_tree, measure = 'auc')

# extract the AUC value
auc_logit@y.values[[1]]
auc_tree@y.values[[1]]


```


#### 5.

```{r}


test_set_t <- test_set %>% 
              mutate(pred_logit_prob = test_logit)

test_set_t <- test_set_t %>%
                mutate(pred_tree_prob = test_tree)


test_set_t <- test_set_t %>%
                mutate(pred_logit_val = case_when(pred_logit_prob >= 0.25 ~ 'Yes',
                                                  pred_logit_prob < 0.25 ~ 'No'))


test_set_t <- test_set_t %>%
                mutate(pred_tree_val = case_when(pred_tree_prob >= 0.25 ~ 'Yes',
                                                  pred_tree_prob < 0.25 ~ 'No'))

test_set_t %>% count(pred_logit_val, Survived) %>% spread(Survived, n)
test_set_t %>% count(pred_tree_val, Survived) %>% spread(Survived, n)


```

