---
title: "COMPSCIX 415.2 Homework 1"
author: "Gokila Rajaiah"
date: "February 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
```

## ggplot Intro (3.2.4):

1. Running ggplot by passing the dataset, will create a blank canvas like below inside which the graph can be plotted by adding layers such as Geom Function etc...

```{r}
ggplot(data = mpg)

```

2. There are 234 rows and 11 columns in the "mpg" dataset.

```{r}

glimpse(mpg)

```


3. "drv" variable tells information about Automobile layout on the vehicle engine and drive wheels. Different categories of them are F - Front Wheel Drive (FWD), R - Rear Wheel Drive (RWD) & 4 = 4 Wheel Drive (4WD)



4. Scatter Plot of Highway Miles Per Gallon vs Number of Cylinders:

    This plot shows that there's a negative relationship between the number of cylinders in the engine and the fuel efficiency. i.e the higher the number of cylinders the lower the highway miles per gallon.

```{r}

ggplot(data = mpg) + geom_point(mapping = aes(x=cyl, y=hwy))

```
5. Scatterplot of class vs drv:

    This plot doesn't give any useful relationship between the two variables class & drv as both are characteristic properties of the car and each class type comes in all the 3 different drive layouts.  

```{r}

ggplot(data = mpg) + geom_point(mapping = aes(x=drv, y=class))

```

## Aesthetic Mappings (3.3.1):


1. Aesthetic is the visual property of the objects in the plot. It can be applied either by mapping the aesthetic property to a variable or by manually setting the aesthetic by name as an argument of geom function. 

  In the below code aesthetic property doesn't show any effect in the plot is because, when using the aesthetic property mapping the aethetic property color should be set to a variable of the data set. Since there's no variable called "blue" in the dataset no effect is seen in the plot. 


```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

```

  In order to the set the point' color to be blue, use the manual setting method of passing the color = "blue" as an arugument to geom_function like below:
  
```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

```

2. Manufacturer, Model, trans, drv, fl and class are categorical variables. displ, cyl, trans, cty, hwy are continuous variables. 

  Running "mpg" shows the data frame with the headers & the types. From the type of the header we can identify if the variable is categorical or continuous. 


```{r}
mpg

```

3. Mapping a continuous variable to Color property will display the variations of values using different saturations of blue, whereas mapping a categorical variable to Color will display distinct values of the property in different colors.


```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x=displ, y=cty, color=hwy))

```
        
   Mapping a continuous variable to size property will map the value to area of the mark. Any ordered continuous variable is suitable to show the variation using size property. 
      

```{r}

ggplot(data=mpg) + geom_point(mapping = aes(x=displ, y=cty, size=cty))

```


  Shapes cannot be mapped using a continous variable as they cannot be ordered. 
      
```{r}

ggplot(data=mpg) + geom_point(mapping = aes(x=displ, y=hwy, shape=drv))

```

4. Mapping same variable to multiple aesthetics will work just fine. This will give more visual cues to differentiate things from having a very slight variation between different values (ex: when color uses two or more colors of same with a sligh variation in the saturation.)


```{r}

ggplot(data=mpg) + geom_point(mapping = aes(x=displ, y=cty, shape=drv, color=drv))

ggplot(data=mpg) + geom_point(mapping = aes(x=displ, y=cty, size=cty, color=cty))

```
5. Stroke aesthetic modifies the width of the border of shape according to the value provided. This can be used with all those shapes with a border, which are the following: 21, 22, 23, 24. Using stroke with shape, colour & fill property you can colour the inside and outside of the shape with different colours. 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y=cty), shape=23, colour = "blue", fill = "white", size = 5, stroke = 2)

```

6. Mapping an aesthetic to an expression will evaluate the expression and uses the result as data to plot. Below is an example where displ < 5 will get evalated to a boolean value and is plotted. 

  Mapping an aesthetic to anything else other than the variable column of property values will just plot the graph by default color without considering the value. I suppose doing this should technically throw an error. 

```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x=trans, y=hwy, color= displ<5))
ggplot(data=mpg) + geom_point(mapping = aes(x=trans, y=hwy, color= 'hello'))

```