---
title: "COMPSCIX 415.2 Homework 1"
author: "Gokila Rajaiah"
date: "February 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## ggplot Intro (3.2.4):

1. Running ggplot by passing the dataset, will create a blank canvas like below inside which the graph can be plotted by adding layers such as Geom Function etc...

```{r}

ggplot(data = mpg)

```

2. There are 234 rows and 11 columns in the "mpg" dataset.

```{r}

glimpse(mpg)

```


3. "drv" variable tells information about Automobile layout on the vehicle engine and drive wheels. Different categories of them are F - Front Wheel Drive (FWD), R - Rear Wheel Drive (RWD) & 4 = 4 Wheel Drive (4WD)



4. Scatter Plot of Highway Miles Per Gallon vs Number of Cylinders:  

  * This plot shows that there's a negative relationship between the number of cylinders in the engine and the fuel efficiency. i.e the higher the number of cylinders the lower the highway miles per gallon.</p>

```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x=cyl, y=hwy))

```

5. Scatterplot of class vs drv:

   * This plot doesn't give any useful relationship between the two variables class & drv as both are characteristic properties of the car and each class type comes in all the 3 different drive layouts.  

```{r fig.cap="Class vs drv", fig.align='center'}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x=drv, y=class))

```

## Aesthetic Mappings (3.3.1):


1. Aesthetic is the visual property of the objects in the plot. It can be applied either by mapping the aesthetic property to a variable or by manually setting the aesthetic by name as an argument of geom function. 

  * In the below code aesthetic property doesn't show any effect in the plot is because, when using the aesthetic property mapping the aethetic property color should be set to a variable of the data set. Since there's no variable called "blue" in the dataset no effect is seen in the plot. 


```{r fig.cap="Displ vs Hwy using incorrect value of aes color", fig.align='center'}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

```

 
 * In order to the set the point' color to be blue, use the manual setting method of passing the color = "blue" as an arugument to geom_function like below:
  
```{r fig.cap="Correct way to manually set the color of points in scatterplot", fig.align='center'}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

```

2. Manufacturer, Model, trans, drv, fl and class are categorical variables. displ, cyl, trans, cty, hwy are continuous variables. 

  * Running "mpg" shows the data frame with the headers & the types. From the type of the header we can identify if the variable is categorical or continuous. 


```{r}
mpg

```

3. Mapping a continuous variable to Color property will display the variations of values using different saturations of blue, whereas mapping a categorical variable to Color will display distinct values of the property in different colors.


```{r fig.cap="Mapping a continuous variable to Color (displ vs cty)", fig.align='center'}

ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=cty, color=cty))

```
        
   * Mapping a continuous variable to size property will map the value to area of the mark. Any ordered continuous variable is suitable to show the variation using size property. 
      

```{r fig.cap="Mapping a continuous variable to Size (displ vs cty)", fig.align='center'}

ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=cty, size=cty))

```


  * Shapes cannot be mapped using a continous variable as they cannot be ordered. 
      
```{r fig.cap="Mapping a discrete variable to Shape (displ vs hwy)", fig.align='center'}

ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=hwy, shape=drv))

```

4. Mapping same variable to multiple aesthetics will work just fine. This will give more visual cues to differentiate things from having a very slight variation between different values (ex: when color uses two or more colors of same with a slight variation in the saturation.)


```{r, fig.cap="Using multiple aesthetics to same variable", fig.align='center'}

ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=cty, shape=drv, color=drv))

ggplot(data=mpg) + 
  geom_point(mapping = aes(x=displ, y=cty, size=cty, color=cty))

```
5. Stroke aesthetic modifies the width of the border of shape according to the value provided. This can be used with all those shapes with a border, which are the following: 21, 22, 23, 24. Using stroke with shape, colour & fill property we can colour the inside and outside of the shape with different colours. 

```{r fig.cap="Using stroke property to geom_points", fig.align='center'}

ggplot(data = mpg) +
  geom_point(mapping = aes(x=displ, y=cty), shape=23, colour = "blue", fill = "white", size = 5, stroke = 2)

```

6. Mapping an aesthetic to an expression will evaluate the expression and uses the result as data to plot. Below is an example where displ < 5 will get evalated to a boolean value and is plotted. 

  * Mapping an aesthetic to anything else other than the variable column of property values will just plot the graph by default color without considering the value. I suppose doing this should technically throw an error. 

```{r fig.cap="Mapping aesthetic to Expression & unknown variable", fig.align='center'}

ggplot(data=mpg) + 
  geom_point(mapping = aes(x=trans, y=hwy, color= displ<5))

ggplot(data=mpg) + 
  geom_point(mapping = aes(x=trans, y=hwy, color= 'hello'))


```

## Facets:

1. nRow & nCol are arguments that can be used to pass the number of rows and columns to use to display the plots in the Facet.  
   * dir is other option that controls the layout of individual panels. This option sets the direction of panel. 
   
   * In Facet_grid, the number of rows & cols are controlled by the number of distinct values in the variables used to plot the grid. 


```{r, fig.cap="ncol & nrow usage in Facet_Wrap", fig.align='center'}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 4, ncol=2)

```

2. When plotting a facet_grid, if the number of rows & cols are different, then the variable with more distinct values should be used as columns and the variable with least distinct values as rows. Scrolling down through the screen to look at plots is not the best way, as the screens are always wider. 


## Geometric Objects

1. Following geom functions are used to create various types of the charts:
  a. geom_line - Line charts
  b. geom_boxplot - Boxplots
  c. geom_histogram - Histograms
  d. geom_area - Area charts
  


```{r, fig.cap="Line chart using geom_line (displ vs hwy)", fig.align='center', message=FALSE, error=FALSE, warning=FALSE}

# creating a dataframe using mpg to include aggregation of hwy by displ. 
dat <- aggregate(mpg, by = list(mpg$displ), mean)
dat <- dat[c(4, 10)]

# using the created dataframe to plot line chart. 
ggplot(data = dat, mapping = aes(x = displ, y = hwy)) +
    geom_line() 
  
```

2. This plot shows the relationship between engine displacement and highway miles for different drv variables. Looking at it shows that 4 wheel drive cars are the ones that provides low mileage and front wheel drive cars gives high mileage performance. 

```{r}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(show.legend = TRUE) + 
  geom_smooth(se = FALSE)

```

3. Show.legend includes the legend associated with that geom' layer. By default if there's any aesthetics that are mapped for a geom layer, legend is included. In the above chart legend is included showing the mark for both point & line.
If show.legend is set to False it will not display the legend associated with that particular layer' aestheic mapping.
Example below shows that both geom layer' legend is set to False. 

```{r, fig.cap="Using show.legend", fig.align='center', warning=FALSE}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(show.legend = TRUE) + 
  geom_smooth(se = FALSE, show.legend = FALSE) 

```
4. "se" argument to geom_smooth is a boolean which takes in input to show confidence interval around the smooth line. 
By default confidence interval level of 0.95 is used. To control the level of CI we can use level argument. Example below shows the CI of level 0.50 around smooth line. we can turn off this by passing False to se. 

```{r, fig.cap="Using se to show CI around smooth line", fig.align='center', Warnig=FALSE}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(show.legend = TRUE) + 
  geom_smooth(level = 0.50)

```

5. Both the code should produce same graphs. Passing in the mapping inside ggplot is same as passing in same mapping in all the geom layers. 

```{r Warnig=FALSE, fig.cap="Passing data inside ggplot vs passing data inside mapping", fig.align='center'}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))

```

6. 

```{r, fig.cap="Code for all 6 graphs in question 3.6.1.6", fig.align='center', warning=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + geom_smooth(se=FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group=drv)) + 
  geom_smooth(se=FALSE) + geom_point()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group=drv, color=drv)) + 
  geom_smooth(se=FALSE) + geom_point()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(se=FALSE, show.legend = FALSE) + geom_point(aes(color=drv))

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(se=FALSE, aes(linetype=drv)) + geom_point(aes(color=drv))


##Below solution found online. 
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(size = 4, colour = "white") + 
  geom_point(aes(colour = drv))

```

## Bar Charts:

There are two types of bar charts. First one is geom_bar which uses the frequency or other aggregation that is provided to mark the height of the bar of each group. Second one is geom_col which uses the exact value in the input data for each group to mark bar height. In other words geom_col uses stat_identity. We can also create geom_col type chart using geom_bar by providing [stat = "identity"] as an argument to aes of geom_bar layer. 

```{r, fig.cap="geom_bar & geom_col to show same graph", fig.align='center'}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))

cnt = count(diamonds, cut)
cnt

ggplot(data = cnt) + 
  geom_col(mapping = aes(x=cut, y=n))

```
