---
title: "COMPSCIX 415.2 Homework 3"
author: "Gokila Rajaiah"
date: "February 17, 2018"
output: html_document
---

```{r}

library(tidyverse)

```

## Position Adjustments (3.8.1):

1. mpg values are rounded when plotting which causes some of the points to overlap which makes it hard to see where most of the data points lies. We can fix this by adding Jitter to the plot which adds some amout of random noise to indicate ovelapping points. Jitters can be added either using Position property of geom_point or as a separate layer of geom_jitter(). 

```{r}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()

```

2. Width & Height controls the amount of horizontal & vertical jittering. The jitter is added in both positive & negative directions so the total spread is twice the value passed. Default is 40% of the resolution of the data. 

```{r}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter(height=50)

```

3. geom_jitter adds random noise to the points to indicate overlapping points. Whereas, geom_count counts the number of obeservations in each point and maps the point area to it. The larger the point is the more number of oversations overlapping. 

```{r fig.cap="geom_count", fig.align='center'}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()

```

4. The default adjustment for geom_boxplot() is position_dodge. 

```{r}

ggplot(data = mpg, mapping = aes(x=class, y=cty, fill=drv)) +
  geom_boxplot(position='dodge') + 
  geom_boxplot(position='identity') 

```

## Coordinate Systems (3.9.1):

2. labs is used to set the title, subtitle & labels for Axis, Legend & Plot. 

```{r}

p <- ggplot(mtcars, aes(mpg, wt, colour = cyl)) + geom_point()
p + labs(colour = "Cylinders")

```


4. At a first glance this plot shows a linear relationship between city mileage & hwy mileage. 

coord_fixed, by default forces a ratio of 1 between the physical representation of data units on the axes.i.e. One unit on x-axis is the same length as one unit on y-axis. This is used whenever both variables have same unit of length, so that they can be compared directly (like apples to apples comparison), which shows a better relationship view in the plot.
Here both axis represents the miles per gallon unit and using fixed scale coordinate system we can see that overall cars have better Highway mileage than City Mileage. 

geom_abline, is used to add a reference line to plot defined by slope & intercept parameters. By default it uses slope 1 & intercept 0. 

Adding both geom_abline & coord_fixed layers to the plot we can see that gas efficiency is on average better for Highway driving & worse for City driving.


```{r}

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()

```

## Workflow Basics (4.4):

1. Second line has a typo. my_variable is declared whereas my_varıable is called, which results in that "variable name not defined error". R is case-sensitive. "Typos matter. Case matters".

2.
  a. "Dota typo" ( We can get "Dota" data but that has to be passed to "data" parameter in ggplot. ;) 
  b. fliter typo. It has to be filter. 
  c. No dataset as "diamond". Typo. It is "diamonds".
  
```{r}


ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)
filter(diamonds, carat > 3)

```


## Data Transformations - Filters (5.2.4):

1.


```{r}

library("nycflights13")

```

```{r}

  (arr_delay_abv_120mins <- filter(flights, arr_delay >= 120))

  (flew_to_houston <- filter(flights, dest == "IAH" | dest == "HOU"))

  (filteredCarriers <- filter(flights, carrier %in% c("UA", "AA", "DL")))

  (departedInSummer <- filter(flights, month %in% c(7, 8, 9)))

  (arrivedTwohrsLate <- filter(flights, arr_delay > 120 && dep_delay <= 0))

  filter(flights, dep_delay >= 60, (dep_delay - arr_delay > 30))
  
  (departedBetMidnightAndSiX <- filter(flights, dep_time >= 2400 | dep_time <= 600 ))


```


2. between() filterting is the shortcut way to use x >= left & x <= right. left & right are boundary values to check for x. 


```{r}

filter(flights, between(dep_time, 600, 2400))

```

3. 8255 flights are missing dep_time. dep_delay, arr_time, arr_delay, tailnum, air_time are other variables missing data. Since both dep_time & arr_time is missing these must be cancelled flights. 

```{r}

sum(is.na(flights$dep_time))

map_dbl(flights, ~sum(is.na(.x)))

filter(flights, is.na(dep_time))

```

4. 

a. NA ^ 0 is not NA because anything to power of 0 is 1. 
b. NA | TRUE is not NA because, if atleast one of the result is TRUE then the expression evaluates to true. In this      second part is TRUE. 

c. FALSE & NA is not NA because atleast one of the two expressions should be tested. In this first one evaluates to FALSE. 


General rule is whenever there is logical expressions involved if atleast one can be evaluated to return a TRUE or FALSE then result should not be NA. 

```{r}

NA ^ 0

NA | TRUE

FALSE & NA

```


## Data Transformations - Select (5.4.1):

1.

```{r}

select(flights, dep_time, dep_delay, arr_time, arr_delay)

select(flights, starts_with("dep"), starts_with("arr"))

select(flights, dep_time, arr_time, ends_with("delay"))

vars <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
select(flights, one_of(vars))

select(flights, matches("^dep|^arr"))

```

2. Including same variable  multiple times will still only include the variable once in the data frame that is returned. 


```{r}

select(flights, dep_time, dep_time, dep_time)

```

3. one_of() selects all variables which matches the variable given in the vectors that is passed. 

```{r}

vars <- c("dep_time", "dep_delay", "arr_time", "arr_delay")
select(flights, one_of(vars))

```

4. Yes. R is case-sensitive. Seeing the select returnig results for contains("TIME") with all those variables which contains "time" doesn't make sense. But, by default each select_helper function ignore the cases of the variables.

To change the default use ignore.case parameter to contains function and set it to False so that cases are taken into account when selecting. 


```{r}

select(flights, contains("TIME"))

select(flights, contains("TIME", ignore.case=F))

```